Edge sign prediction in general graphs and Correlation Clustering

### 2017-11-28 cc/bias_balance.tex
add missing labels
typos and stability in cc

From a narrow practical point of view, balance theories specify which sign assignments of a triangle
are possible. By forbidding triangles with a single negative edge, weak balance gives rise to a
$K$-\emph{consistent} clusters structure on complete graphs. After making these notions more precise
in the rest of this section, we conclude by precisely exposing our learning bias.

@@

We just saw that if a complete graphs is strongly balanced, then it admits a $2$-consistent
clustering (\autoref{thm:structural} proves the converse is true). It is natural to ask under which
conditions a complete signed graph admits a $K>2$ consistent clustering. This is where the notion of
weak balance comes handy. It was noted by
sociologists that among triangles with odd number of negative edges, one is more stable and common
than that the other. Therefore, \textcite{davis1967clustering} relax the strong balance into the
weak balance by considering triangles with three negative edges to be balanced, as illustrated on
\autoref{fig:cc_balance}. Formally, we can modify our two previous definitions of balance to arrive
at weak balance:
\begin{definition}[Weak balance]\label{def:weak_balance}
	A triangle is weakly balanced if it has zero, two or three negative signs. Otherwise it is
	unbalanced. \\
	A complete signed graph is weakly balanced if all its triangles are weakly balanced.
\end{definition}

@@

In real data though, we do not expect either strong or weak balance to hold, for they are fairly
demanding model. Indeed, three of the real networks considered in \autoref{chap:troll} have been
repeatedly shown to be unbalanced, although the extent of this unbalance depends on the importance
given to longer cycles compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}%
{measureUnbalance17}.
% there are also \url{https://arxiv.org/abs/1509.04037} about measuring balance (or frustration
% index) which they claimed has been peer reviewed?
% For the same reason, in virtually any real instances of the \pcc{} problem, the optimal solution
% will incur some disagreements. 
\marginpars{The proof makes essential use of the strong balance (which is ok since next section also
deal with 2-CC) and the completeness of the graph (which is a far less realistic assumption) yet
it would be interesting to see if it can be extended.}
\Textcite[Section 5.5]{BookKleinberg2010}
nonetheless prove that even when only a fraction of the triangle are strongly balanced in a complete
graph, the two clusters structure is still present, although it does not cover the whole graph
any more.

@@

Using these results on weak balance, we can now make our new bias explicit. Drawing parallel with
the previous generative model, we assume that each node $u$ is endowed with an integer $\cluster(u)
\in \{1, \ldots, K\}$ that specifies its cluster and we let the undirected edge $u,v$ be positive if
$\cluster(u) = \cluster(v)$ and negative otherwise. In other words, this corresponds to picking an
arbitrary partition of $V$ and setting the signs of $E$ in such a way that the partition becomes a
consistent clustering. This can still be seen as a generative process because of the initial integer
drawn \uar{},\Todo{Although I'm not sure how I would write this likelihood to be fair} but one that
avoids the two pitfalls discussed before. Indeed, there is no decision to be made by nodes, since
the previous probability has been replaced by a binary function (either both nodes are in the same
cluster or not). Furthermore, since the clustering function is symmetric, the model is inherently
suited to undirected graphs.

>>>>

From a narrow practical point of view, balance theories specify which sign assignments of a triangle
are possible. By forbidding triangles with a single negative edge, weak balance gives rise to a
$K$-\emph{consistent} clusters structure on complete graphs. After making these notions more detailed
in the rest of this section, we conclude by precisely exposing our learning bias.

@@

We just saw that if a complete graph is strongly balanced, then it admits a $2$-consistent
clustering (\autoref{thm:structural} proves that the converse is true). It is natural to ask under which
conditions a complete signed graph admits a $K>2$ consistent clustering. This is where the notion of
weak balance comes handy. It was noted by
sociologists that among triangles with odd number of negative edges, one is more stable and common
than that the other. Therefore, \textcite{davis1967clustering} relax the strong balance into the
weak balance by considering triangles with three negative edges to be balanced, as illustrated on
\autoref{fig:cc_balance}. Formally, we can modify our two previous definitions of balance to arrive
at weak balance:
\begin{definition}[Weak balance]\label{def:weak_balance}
	A triangle is weakly balanced if it has zero, two or three negative signs. Otherwise it is
	unbalanced. \\
	A complete signed graph is weakly balanced if all its triangles are weakly balanced.
\end{definition}

@@

In real data though, we do not expect either strong or weak balance to hold, for they are fairly
demanding model. Indeed, three of the real networks considered in \autoref{chap:troll} have been
repeatedly shown to be unbalanced, although the extent of this unbalance depends on the importance
given to longer cycles compared with triads~\autocites{Facchetti2011isingmodel}{measureUnbalance14}%
{measureUnbalance17}.
% there are also \url{https://arxiv.org/abs/1509.04037} about measuring balance (or frustration
% index) which they claimed has been peer reviewed?
% For the same reason, in virtually any real instances of the \pcc{} problem, the optimal solution
% will incur some disagreements. 
%\marginpars{The proof makes essential use of the strong balance (which is ok since next section also
%deal with 2-CC) and the completeness of the graph (which is a far less realistic assumption) yet
%it would be interesting to see if it can be extended.}
\Textcite[Section 5.5]{BookKleinberg2010}
nonetheless prove that even when only a fraction of the triangle are strongly balanced in a complete
graph, the two clusters structure is still present, although it does not cover the whole graph
any more.

@@

\label{text:cc_new_bias}
Using these results on weak balance, we can now make our new bias explicit. Drawing parallel with
the previous generative model, we assume that each node $u$ is endowed with an integer $\cluster(u)
\in \{1, \ldots, K\}$ that specifies its cluster and we let the undirected edge $u,v$ be positive if
$\cluster(u) = \cluster(v)$ and negative otherwise. In other words, this corresponds to picking an
arbitrary partition of $V$ and setting the signs of $E$ in such a way that the partition becomes a
consistent clustering. This can still be seen as a generative process because of the initial integer
drawn \uar{}, but one that
%\Todo{Although I'm not sure how I would write this likelihood to be fair}
avoids the two pitfalls discussed before. Indeed, there is no decision to be made by nodes, since
the previous probability has been replaced by a binary function (either both nodes are in the same
cluster or not). Furthermore, since the clustering function is symmetric, the model is inherently
suited to undirected graphs.

### 2017-11-28 cc/cc_non_worst_case.tex
typos and stability in cc

\Textcite{Joachims2005} analyze a more refined weighted model, where weights are generated by a
probability distribution whose mean on true positive edges is larger than $\mu^+>0$ and whose mean
on true negative edges is smaller than $\mu^-<0$. They give a finite-sample bound on the number of
node misclustered w.r.t the planted partition as a function on the probability distribution
parameters. Indeed, as pointed out by \textcite{plantedAilon09}, independent and uniform noise is
not a good model of real situations, where the input of \pcc{}, \ie{} the similarity between nodes,
is often the result of a preprocessing, which may present strong correlations. Therefore, instead of
measuring the quality of a solution against in the input (\ie{} the similarity information), they
argue it is more sensible to measure it against the (unknown) true optimal clustering that gave rise
to the input, and show that \ccpivot{} allows that thanks to a new analysis. \Textcite{Mathieu2010}
also consider an adversarial model, where all edges are flipped with probability $p$ but the
adversary then decide whether to show us the true sign or the flipped sign. On complete unweighted
graphs, they find a solution of \mind{} at most $1 + O(n^{-\nicefrac{1}{6}})$ times the optimal
whenever $p \leq \shalf - n^{-\nicefrac{1}{3}}$ by rounding the usual SDP solution. If in addition
$p\leq\nicefrac{1}{3}$, there is no adversary and each planted cluster has at least
$\Theta(\sqrt{n})$ nodes, then the planted partition can be recovered exactly.
\Textcite{Makarychev2014} study the same adversarial model but on general weighted graphs, giving a PTAS
for \mind{} when $p\leq \nicefrac{1}{4}$. Under additional assumptions on the density of edges, they
present another algorithm that finds the ground truth clustering with an arbitrarily small
classification error.\footnote{They also mention a potentially interesting related
work~\autocite{unobservedCC14}.} % http://www.jmlr.org/papers/v15/chen14a.html
% \textcite{Noisy2CCLocality16} study the recovery of an arbitrary planted 2-partition under noisy
% measurement in special kind of graphs where each node is connected with a local neighborhood and
% show computationally efficient algorithms that only query an information theoretically optimal
% number of pair. I think it was extended in \url{https://doi.org/10.1109/TIT.2016.2600566}

@@

Whereas clustering objective functions are \NPh{} to optimize, we expect that meaningful instances
in which we are interested have additional structure which allows for guaranteed polynomial time
algorithms, see for instance \autocite{clusteringFeasibility15} for a critical overview of some
proposed such notions of structure. Informally, a general idea is that the clustering should not
change (or at least very little) if the data are slightly perturbed. For instance, a weighted graph
is \emph{$\alpha$-stable} (with $\alpha>1$) for some partition objective if its optimal partition
remains the same whenever every weight $w_i$ is multiplied by a factor $c_i$ between $1$ and
$\alpha$.
Finally, we say that an algorithm is \emph{robust} if, given an instance $\mathcal{I}$ and
in polynomial time, it: returns the optimal solution of $\mathcal{I}$ if $\mathcal{I}$ is
$\alpha$-stable; and if $\mathcal{I}$ is not $\alpha$-stable, either returns the optimal solution of
$\mathcal{I}$ or reports that $\mathcal{I}$ is not $\alpha$-stable. This is a handy property, as in
general, we cannot check the stability of an instance in practice as we do not know the optimal
solution hence we cannot tell whether it changes or not under perturbations.

See \autocite{StableCC09} for application to \mind{} on complete graphs and \autocite{StableLP09}
for more general study of the solution stability of the canonical \mind{} LP in any kind of graph.

>>>>

\Textcite{Joachims2005} analyze a more refined weighted model, where weights are generated by a
probability distribution whose mean on true positive edges is larger than $\mu^+>0$ and whose mean
on true negative edges is smaller than $\mu^-<0$. They give a finite-sample bound on the number of
nodes misclustered w.r.t the planted partition as a function on the probability distribution
parameters. Indeed, as pointed out by \textcite{plantedAilon09}, independent and uniform noise is
not a good model of real situations, where the input of \pcc{}, \ie{} the similarity between nodes,
is often the result of a preprocessing, which may present strong correlations. Therefore, instead of
measuring the quality of a solution against the input (\ie{} the similarity information), they
argue it is more sensible to measure it against the (unknown) true optimal clustering that gave rise
to the input, and show that \ccpivot{} allows this, thanks to a new analysis. \Textcite{Mathieu2010}
also consider an adversarial model, where all edges are flipped with probability $p$ but the
adversary then decides whether to reveal the true sign or the flipped sign. On complete unweighted
graphs, they find a solution of \mind{} at most $1 + O(n^{-\nicefrac{1}{6}})$ times the optimal
whenever $p \leq \shalf - n^{-\nicefrac{1}{3}}$ by rounding the usual SDP solution. If, in addition,
there is no adversary, $p\leq\nicefrac{1}{3}$ and each planted cluster has at least
$\Theta(\sqrt{n})$ nodes, then the planted partition can be recovered exactly.
\Textcite{Makarychev2014} study the same adversarial model but on general weighted graphs, giving a PTAS
for \mind{} when $p\leq \nicefrac{1}{4}$. Under additional assumptions on the density of edges, they
present another algorithm that finds the ground truth clustering with an arbitrarily small
classification error.
% \footnote{They also mention a potentially interesting related work~\autocite{unobservedCC14}.}
% http://www.jmlr.org/papers/v15/chen14a.html
% \textcite{Noisy2CCLocality16} study the recovery of an arbitrary planted 2-partition under noisy
% measurement in special kind of graphs where each node is connected with a local neighborhood and
% show computationally efficient algorithms that only query an information theoretically optimal
% number of pair. I think it was extended in \url{https://doi.org/10.1109/TIT.2016.2600566}

@@

Whereas clustering objective functions are \NPh{} to optimize, we expect meaningful instances
in which we are interested to have additional structure which allows for guaranteed polynomial time
algorithms. We refer the reader to a critical overview~\autocite{clusteringFeasibility15} of some
such notions of structure proposed recently. Informally, a general idea is that the clustering should not
change (or at least very little) if the data are slightly perturbed. For instance, a weighted graph
is \emph{$\alpha$-stable} (with $\alpha>1$) for some partition objective if its optimal partition
remains the same whenever every weight $w_i$ is multiplied by a factor $c_i$ between $1$ and
$\alpha$.
Another notion is the $(c, \epsilon)$-approximation stability. Formally, a dataset $X$ is $(c,
\epsilon)$-stable if, for an objective function $\Phi$, any clustering whose cost is within a factor
$c$ of the optimal cost $OPT_{\Phi}$ is $\epsilon$-close to the optimal clustering (as measured by
the fraction of points in which the two clustering disagree). When the objective is \mind{} on
complete graphs, \textcite{StableCC09} show that for $(1+\alpha, \epsilon)$-instances, the
approximation algorithms we describe in \autoref{ssub:cc_harness_approx} find a solution that is
$(\nicefrac{49}{\alpha} + 1)\epsilon$ close to the optimal clustering.

It is also possible to study the stability of a \pcc{} instance over a general graph w.r.t edge
weights via its canonical \mind{} LP. \Textcite{StableLP09} present a method to determine to which
extent the weights of a \pcc{} instance can be perturbed before the optimal solution changes.

### 2017-11-28 cc/cc_problem_and_esp.tex
typos and stability in cc

In \pcc{}, we formalize this problem by considering objects as the nodes of a graph $G$, whose edge
weights encode similarity. Namely, in the most general case, for nodes $u$ and $v$, the edge between
$u$ and $v$ is associated with two positive numbers:
$w_{u,v}^+$ denotes the strength of the similarity between $u$ and $v$;
$w_{u,v}^-$ denotes the strength of the dissimilarity between $u$ and $v$.
Note, however, that in many applications, only one of these two numbers is nonzero, in which case
we more conveniently set $w_{u,v} = \begin{cases}
	 w_{u,v}^+ & \quad \text{if } w_{u,v}^+ > 0 \text{ and } w_{u,v}^-=0 \\
	-w_{u,v}^- & \quad \text{if } w_{u,v}^+ = 0 \text{ and } w_{u,v}^->0 \\
\end{cases}$

@@

\esp{} can clearly be casted as a supervised classification problem where, given a labelled training
set of $m_0$ edges $\mathcal{X}=\{(e_1, y_1), \ldots, (e_{m_0}, y_{m_0})\}$, we can extract for
every edge a feature vector in $\Rbb^d$, pick an hypothesis class like linear models, find among
this class the hypothesis minimizing the empirical risk, and use it to predict the sign of edges in
the testing set. In contrast, \pcc{} in an agnostic problem. More precisely, the hypothesis class is
fixed, as it consists of all possible partitions of the nodeset. We are trying to find the partition
that best approximates the observed signs, but each weakly unbalanced cycle is gonna cost us at
least one unavoidable mistakes. Despite this fundamental difference in their nature, \pcc{} is
connected with \esp{} at two levels, theoretical and practical. On the theoretical side, computing
the minimum number of disagreements gives a measure of the \esp{} problem complexity.  Intuitively,
given a labelled subgraph $H$ of $G$ (that is the training set), if there are few disagreements, the
\enquote{quasi} connected components of $H^+$ should give a reasonable estimation of the underlying
consistent clustering. Predicting the signs based on this clustering should results in few errors.
More formally, one can show that by finding a clustering $\cluster^\star$ of $H$ that has less
disagreements than $\kappa$ times the smallest number of disagreements $\Delta$, the number of
errors on the testing set will be at most of order $\kappa\Delta + \sqrt{mn\log
n}$~\autocite[Theorem 6]{Cesa-Bianchi2012b}. These theoretical considerations naturally lead to
practical batch algorithm, which first finds a clustering minimizing as much as possible the number
of disagreements on the training set and then predicts signs according to this clustering. Our
assumption in doing so is that, according to our bias, our graph has few disagreements. Therefore,
provided we can find a clustering close to the optimal, we will not make too many prediction
mistakes.

@@

Furthermore, as mentioned in the introduction of this thesis, \pcc{} has also been used in several
domains. For instance in computer vision, it is used to segment images in
2D~\autocites{Kim2011}{Bagon2011}{CellSeg14} or 3D~\autocites{VolumeSegmentation12}{Beier2015}, to
simplify 3D shapes~\autocite{Shape3D17} and to track target across sequential video
frames~\autocite{multiTracking15}. When the nodes are words, \pcc{} is employed to identify
coreference~\autocites{graphicalCoreference04}{Elsner2009} or cluster synonym
words~\autocite{SignedWordRatings}. It plays a large role in biology, to clusters
genes~\autocite{Ben-Dor99}, identify mutation regions in chromosomes~\autocite{Das2015} or finding
stable subsystems~\autocite{monotoneBiology07}. Applied to social networks, it helps analyze
political assemblies~\autocites{Mendonca2015}{Jiang2015}{BrazilCC17} and international
relations~\autocites{Traag2009}{CommunityUN12}. Finally, given large databases with duplicated
records, it naturally models entity
resolution~\autocites{Crosslingual07}{DeDup09}{LargeScaleDeDup09}{WeightedER14}.

>>>>

In \pcc{}, we formalize this problem by considering objects as the nodes of a graph $G$, whose edge
weights encode similarity. Namely, in the most general case, for nodes $u$ and $v$, the edge between
$u$ and $v$ is associated with two positive numbers:
$w_{u,v}^+$ denotes the strength of the similarity between $u$ and $v$, while
$w_{u,v}^-$ denotes the strength of the dissimilarity between $u$ and $v$.
Note, however, that in many applications, only one of these two numbers is nonzero, in which case
we more conveniently set $w_{u,v} = \begin{cases}
	 w_{u,v}^+ & \quad \text{if } w_{u,v}^+ > 0 \text{ and } w_{u,v}^-=0 \\
	-w_{u,v}^- & \quad \text{if } w_{u,v}^+ = 0 \text{ and } w_{u,v}^->0 \\
\end{cases}$

@@

\esp{} can clearly be casted as a supervised classification problem where, given a labelled training
set of $m_0$ edges $\mathcal{X}=\{(e_1, y_1), \ldots, (e_{m_0}, y_{m_0})\}$, we can extract for
every edge a feature vector in $\Rbb^d$, pick an hypothesis class like linear models, find among
this class the hypothesis minimizing the empirical risk, and use it to predict the sign of edges in
the testing set. In contrast, \pcc{} in an agnostic problem. More precisely, the hypothesis class is
fixed, as it consists of all possible partitions of the nodeset. We are trying to find the partition
that best approximates the observed signs, but each weakly unbalanced cycle is gonna cost us at
least one unavoidable mistake. Despite this fundamental difference in their nature, \pcc{} is
connected with \esp{} at two levels, theoretical and practical. On the theoretical side, computing
the minimum number of disagreements gives a measure of the \esp{} problem complexity.  Intuitively,
given a labelled subgraph $H$ of $G$ (that is the training set), if there are few disagreements, the
\enquote{quasi} connected components of $H^+$ should give a reasonable estimation of the underlying
consistent clustering. Predicting the signs based on this clustering should results in few errors.
More formally, one can show that by finding a clustering $\cluster^\star$ of $H$ that has less
disagreements than $\kappa$ times the smallest number of disagreements $\Delta$, the number of
errors on the testing set will be at most of order $\kappa\Delta + \sqrt{mn\log
n}$~\autocite[Theorem 6]{Cesa-Bianchi2012b}. These theoretical considerations naturally lead to
practical a batch algorithm, which first finds a clustering minimizing as much as possible the number
of disagreements on the training set and then predicts signs according to this clustering. Our
assumption in doing so is that, according to our bias, our graph has few disagreements. Therefore,
provided we can find a clustering close to the optimal, we will not make too many prediction
mistakes.

@@

Furthermore, as mentioned in the introduction of this thesis \vpageref{sub:intro_signed_graphs},
\pcc{} have been used in many
domains. For instance in computer vision, it is used to segment images in
2D~\autocites{Kim2011}{Bagon2011}{CellSeg14} or 3D~\autocites{VolumeSegmentation12}{Beier2015}, to
simplify 3D shapes~\autocite{Shape3D17} and to track target across sequential video
frames~\autocite{multiTracking15}. When the nodes are words, \pcc{} is employed to identify
coreference~\autocites{graphicalCoreference04}{Elsner2009} or cluster synonym
words~\autocite{SignedWordRatings}. It plays a large role in biology, to clusters
genes~\autocite{Ben-Dor99}, identify mutation regions in chromosomes~\autocite{Das2015} or finding
stable subsystems~\autocite{monotoneBiology07}. Applied to social networks, it helps analyze
political assemblies~\autocites{Mendonca2015}{Jiang2015}{BrazilCC17} and international
relations~\autocites{Traag2009}{CommunityUN12}. Finally, given large databases with duplicated
records, it naturally models entity
resolution~\autocites{Crosslingual07}{DeDup09}{LargeScaleDeDup09}{WeightedER14}.
